canaccessR: An open data product for transit accessibility analysis in Canada's largest metropolitan areas.

### 2025-05-03 paper/Manuscript-Data-Package-cover.Rmd
Fixed some errors for submission to EPB

---
title: "canaccessR: An open data product for transit accessibility analysis in Canada’s largest metropolitan areas."
author:
  - name: João Pedro Figueira Amorim Parga
    email: joao.parga@mail.utoronto.ca
    affil-id: 1
    orcid: 0000-0002-4105-5927
    footnote: 1
  - name: Anastasia Soukhov
    email: soukhoa@mcmaster.ca
    affil-id: 2
    orcid: 0000-0003-4371-4831
  - name: Robert Nutifafa Arku
    email: rn.arku@mail.utoronto.ca
    affil-id: 1
    orcid: 0000-0002-2018-886X
  - name: Christopher Higgins
    email: cd.higgins@utoronto.ca
    affil-id: 3
    orcid: 0000-0002-3551-7750
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affil-id: 2
    orcid: 0000-0001-6912-9919
affiliations:
  - id: 1
    name: University of Toronto
  - id: 2
    name: McMaster University
  - id: 3
    name: University of Toronto Scarborough
# address:
#   - code: CODE
#     address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "In this paper, we describe the {canaccessR} package, an open data product (ODP) created in R that contains public transit travel time estimates to employment locations and grocery stores across Canada's 12 largest metropolitan areas. We calculate travel time matrices (TTM) from and to each Dissemination Area (DA) within these regions for the years 2019 and 2023. We add value to the urban analytics community by processing and integrating raw data, and disseminating user-ready data in the domain of transportation accessibility in Canada. To do so, we use the {r5r} R package, General Transit Feed Specification (GTFS), OpenStreetMap (OSM), DMTI’s Enhanced Points of Interest, and Statistics Canada Census data. This data package can be used by researchers, practitioners, and transit agencies to estimate accessibility levels to these two essential destinations within these urban areas. Moreover, travel time matrices are computed from DA centroid to DA centroid, which means that they can be adapted for use in applications with any type of destination that is aggregated at the DA level. Finally, as an ODP, the {canaccess} package allows for open exploration, use, and contribution by users through its GitHub repository."

>>>>

---
title: "canaccessR: An open data product for transit accessibility analysis in Canada’s largest metropolitan areas."
author:
  - name: João Figueroa Amorim PARGA
    email: joao.parga@mail.utoronto.ca
    affil-id: 1
    orcid: 0000-0002-4105-5927
    footnote: 1
  - name: Anastasia SOUKHOV
    email: soukhoa@mcmaster.ca
    affil-id: 2
    orcid: 0000-0003-4371-4831
  - name: Robert Nutifafa ARKU
    email: rn.arku@mail.utoronto.ca
    affil-id: 1
    orcid: 0000-0002-2018-886X
  - name: Christopher HIGGINS
    email: cd.higgins@utoronto.ca
    affil-id: 3
    orcid: 0000-0002-3551-7750
  - name: Antonio PÁEZ
    email: paezha@mcmaster.ca
    affil-id: 2
    orcid: 0000-0001-6912-9919
affiliations:
  - id: 1
    name: University of Toronto
  - id: 2
    name: McMaster University
  - id: 3
    name: University of Toronto Scarborough
# address:
#   - code: CODE
#     address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "In this paper, we describe the {canaccessR} package, an open data product (ODP) created in R that contains public transit travel time estimates to employment locations and grocery stores across Canada's 12 largest metropolitan areas. We calculate travel time matrices (TTM) from and to each Dissemination Area (DA) within these regions for the years 2019 and 2023. We add value to the urban analytics community by processing and integrating raw data, and disseminating user-ready data in the domain of transportation accessibility in Canada. To do so, we use the {r5r} R package, General Transit Feed Specification (GTFS), OpenStreetMap (OSM), DMTI’s Enhanced Points of Interest, and Statistics Canada Census data. This data package can be used by researchers, practitioners, and transit agencies to estimate accessibility levels to these two essential destinations within these urban areas. Moreover, travel time matrices are computed from DA centroid to DA centroid, which means that they can be adapted for use in applications with any type of destination that is aggregated at the DA level. Finally, as an ODP, the {canaccess} package allows for open exploration, use, and contribution by users through its GitHub repository."

### 2025-05-03 paper/Manuscript-Data-Package-cover.tex
Fixed some errors for submission to EPB

\author{João Pedro Figueira Amorim Parga\affilnum{}, Anastasia
Soukhov\affilnum{}, Robert Nutifafa Arku\affilnum{}, Christopher
Higgins\affilnum{}, Antonio Páez\affilnum{}}

>>>>

\author{João Figueroa Amorim PARGA\affilnum{}, Anastasia
SOUKHOV\affilnum{}, Robert Nutifafa ARKU\affilnum{}, Christopher
HIGGINS\affilnum{}, Antonio PÁEZ\affilnum{}}

### 2025-05-03 paper/Manuscript-Data-Package.Rmd
Fixed some errors for submission to EPB

keywords: Public transit accessibility; open data products (ODPs); R data package; travel time matrices.
classoption:
  - Royal
  - times
bibliography: bibfile
bibliographystyle: sageh
output:
  #rticles::sage_article:
  #  keep_tex: yes
  #  fig_caption: yes
  word_document
header-includes: \usepackage{float}
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages
library(canaccessR)
library(data.table)
library(glue)
library(gt)
library(gtsummary)
# library(summarytools)
# library(sf)
library(tidyverse)
# library(tmap)
# library(tmaptools)
```

@@

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. Table @tbl-table_1 summarizes the travel time estimates. The travel time objects contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DAs in the sample and 18,519,897 pairs from population to grocery stores. Considering all areas combined, the mean travel time to jobs was 78 and 76 to grocery stores.

```{r, fig.pos="H"}
#| label: tbl-table_1
#| tbl-cap: "Descriptive statistics of the travel time matrices from the {canaccessR} package."
#| echo: false
#| message: false

sk_full %>% 
  gt() %>% 
  fmt_number(use_seps = TRUE, decimals = FALSE) |>  
  tab_options( 
    table.font.size = 10,
    table.width = pct(100) 
    )

@@

This section presents a usage example of the package through a visual representation of its data. In Figure @fig-travel_time_emp_grc_plot, we show the spatial distribution of the travel time matrices for the metropolitan region of Montréal. In it, we see the median travel time from each DA to employment (left) and to grocery stores (right). The plot shows that moving away from the city core increases the necessary travel time by public transit to reach both employment locations and grocery stores. This result is expected for employment, given the typical concentration of this type of opportunity in downtown centers, but it also points out at a distribution of grocery stores that also requires increased travel for non-central city residents.

```{r, fig.pos="H"}
#| label: fig-travel_time_emp_grc_plot
#| fig-cap: "Estimated median travel time (minutes) per Dissemination Area to jobs (left) and grocery stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021), and DA and planning boundaries of the Montréal metropolitan area use Statistics Canada data (Statistics Canada, 2016)."
#| out-width: 100%
#| echo: false
knitr::include_graphics("../figures/patch_tt_emp_grc.jpg")
```
<!--AS: Where is the 0-30 bin for grocery stores? It would be helpful to change the plot so it reads '0-30' '30-60', '60-90', 90-120min -->

A more thorough example of the package's use can be found in the School of Cities' recent report on Canada's Urban Infrastructure Deficit. In its 11th Chapter, we use the travel time matrices to estimate accessibility metrics to jobs and grocery stores before and after the pandemic [@pargaDemocraticAccessOur2024]. We then compare how changes affected groups differently according to their spatial distribution and income level, thus making explicit the connection between the package's information and matters of equity in transportation. The report is freely available for download at the State of Cities Summit [website](https://stateofcitiessummit.ca/report).

>>>>

keywords: Public transit accessibility; open data products (ODPs); R data package; travel time matrices.
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
   keep_tex: yes
   fig_caption: yes
#  word_document
header-includes: \usepackage{float}
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages
library(canaccessR)
library(data.table)
library(glue)
#library(gt)
library(gtsummary)
library(kableExtra)
# library(summarytools)
# library(sf)
library(tidyverse)
# library(tmap)
# library(tmaptools)
```

@@

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. Table \ref{tab:table_1} summarizes the travel time estimates. The travel time objects contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DAs in the sample and 18,519,897 pairs from population to grocery stores. Considering all areas combined, the mean travel time to jobs was 78 and 76 to grocery stores.

```{r, fig.pos="H"}
#| echo: false
#| message: false

sk_full %>%
  kableExtra::kable("latex",
                    booktabs = TRUE,
                    digits = 2,
                    caption = "\\label{tab:table_1}Descriptive statistics of the travel time matrices from the {canaccessR} package.") %>%
kable_styling(latex_options = c("striped", "scale_down"))

@@

This section presents a usage example of the package through a visual representation of its data. In Figure \ref{fig:fig-travel_time_emp_grc_plot}, we show the spatial distribution of the travel time matrices for the metropolitan region of Montréal. In it, we see the median travel time from each DA to employment (left) and to grocery stores (right). The plot shows that moving away from the city core increases the necessary travel time by public transit to reach both employment locations and grocery stores. This result is expected for employment, given the typical concentration of this type of opportunity in downtown centers, but it also points out at a distribution of grocery stores that also requires increased travel for non-central city residents.

```{r, fig.pos="H", fig.cap={"\\label{fig:fig-travel_time_emp_grc_plot}Estimated median travel time (minutes) per Dissemination Area to jobs (left) and grocery stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021), and DA and planning boundaries of the Montréal metropolitan area use Statistics Canada data (Statistics Canada, 2016)."}}
##| label: fig-travel_time_emp_grc_plot
##| fig-cap: "Estimated median travel time (minutes) per Dissemination Area to jobs (left) and grocery stores (right). Public transit travel times are #calculated using {r5r} (Pereira et al., 2021), and DA and planning boundaries of the Montréal metropolitan area use Statistics Canada data (Statistics #Canada, 2016)."
#| out-width: 100%
#| echo: false
knitr::include_graphics("../figures/patch_tt_emp_grc.jpg")
```
<!--AS: Where is the 0-30 bin for grocery stores? It would be helpful to change the plot so it reads '0-30' '30-60', '60-90', 90-120min -->

A more thorough example of the package's use can be found in the School of Cities' recent report on Canada's Urban Infrastructure Deficit. In its 11th Chapter, the travel time matrices in the package are used to estimate accessibility metrics to jobs and grocery stores before and after the pandemic [@pargaDemocraticAccessOur2024]. This work also compares how changes affected groups differently according to their spatial distribution and income level, thus making explicit the connection between the package's information and matters of equity in transportation. The report is freely available for download at the State of Cities Summit [website](https://stateofcitiessummit.ca/report).

### 2025-05-03 paper/Manuscript-Data-Package.tex
Fixed some errors for submission to EPB

\usepackage{float}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}

@@

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. Table \citet{tbl-table_1}
summarizes the travel time estimates. The travel time objects contain,
in total, 97,784,850 origin-destination pairs (observations) from
population to employment locations for all the DAs in the sample and
18,519,897 pairs from population to grocery stores. Considering all
areas combined, the mean travel time to jobs was 78 and 76 to grocery
stores.

\begin{table}[t]
\fontsize{7.5pt}{9.0pt}\selectfont
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}clrrrrrr}
\toprule
Study Region Name & Destination & Observations & Mean & Sd & P25 & P50 & P75 \\ 
\midrule\addlinespace[2.5pt]
All regions & Employment & 97,784,850 & 78 & 25 & 59 & 80 & 99 \\ 
Toronto & Employment & 41,038,062 & 82 & 25 & 64 & 85 & 103 \\ 
Montréal & Employment & 30,091,411 & 76 & 25 & 57 & 77 & 97 \\ 
Vancouver & Employment & 12,254,478 & 76 & 25 & 58 & 77 & 96 \\ 
Calgary & Employment & 3,022,244 & 75 & 23 & 59 & 75 & 91 \\ 
Ottawa & Employment & 2,936,912 & 74 & 24 & 56 & 74 & 93 \\ 
Edmonton & Employment & 2,057,486 & 70 & 23 & 53 & 70 & 87 \\ 
Québec City & Employment & 1,473,702 & 72 & 25 & 54 & 72 & 93 \\ 
Winnipeg & Employment & 1,574,205 & 62 & 22 & 46 & 61 & 76 \\ 
Hamilton & Employment & 2,089,172 & 82 & 28 & 61 & 87 & 106 \\ 
Waterloo & Employment & 552,826 & 71 & 27 & 49 & 69 & 93 \\ 
London & Employment & 415,174 & 61 & 21 & 46 & 60 & 74 \\ 
Halifax & Employment & 279,178 & 66 & 25 & 47 & 65 & 84 \\ 
All regions & Groceries Stores & 18,519,897 & 76 & 26 & 56 & 76 & 97 \\ 
Toronto & Groceries Stores & 8,512,874 & 80 & 25 & 61 & 82 & 101 \\ 
Montréal & Groceries Stores & 2,993,965 & 73 & 26 & 53 & 74 & 94 \\ 
Vancouver & Groceries Stores & 4,540,106 & 73 & 25 & 53 & 74 & 93 \\ 
Calgary & Groceries Stores & 502,757 & 70 & 22 & 54 & 70 & 86 \\ 
Ottawa & Groceries Stores & 619,791 & 72 & 24 & 55 & 72 & 91 \\ 
Edmonton & Groceries Stores & 324,030 & 68 & 23 & 51 & 67 & 84 \\ 
Québec City & Groceries Stores & 234,600 & 70 & 26 & 51 & 70 & 89 \\ 
Winnipeg & Groceries Stores & 362,566 & 58 & 21 & 43 & 57 & 72 \\ 
Hamilton & Groceries Stores & 281,118 & 84 & 29 & 60 & 91 & 110 \\ 
Waterloo & Groceries Stores & 44,726 & 66 & 28 & 44 & 62 & 90 \\ 
London & Groceries Stores & 52,617 & 55 & 19 & 42 & 55 & 67 \\ 
Halifax & Groceries Stores & 50,747 & 63 & 26 & 42 & 63 & 83 \\ 
\bottomrule
\end{tabular*}
\end{table}

@@

This section presents a usage example of the package through a visual
representation of its data. In Figure
\citet{fig-travel_time_emp_grc_plot}, we show the spatial distribution
of the travel time matrices for the metropolitan region of Montréal. In
it, we see the median travel time from each DA to employment (left) and
to grocery stores (right). The plot shows that moving away from the city
core increases the necessary travel time by public transit to reach both
employment locations and grocery stores. This result is expected for
employment, given the typical concentration of this type of opportunity
in downtown centers, but it also points out at a distribution of grocery
stores that also requires increased travel for non-central city
residents.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{../figures/patch_tt_emp_grc} \caption{Estimated median travel time (minutes) per Dissemination Area to jobs (left) and grocery stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021), and DA and planning boundaries of the Montréal metropolitan area use Statistics Canada data (Statistics Canada, 2016).}\label{fig:fig-travel_time_emp_grc_plot}
\end{figure}

A more thorough example of the package's use can be found in the School
of Cities' recent report on Canada's Urban Infrastructure Deficit. In
its 11th Chapter, we use the travel time matrices to estimate
accessibility metrics to jobs and grocery stores before and after the
pandemic \citep{pargaDemocraticAccessOur2024}. We then compare how
changes affected groups differently according to their spatial
distribution and income level, thus making explicit the connection
between the package's information and matters of equity in
transportation. The report is freely available for download at the State
of Cities Summit \href{https://stateofcitiessummit.ca/report}{website}.

@@

\bibliographystyle{sageh}
\bibliography{bibfile}

>>>>

\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. Table \ref{tab:table_1}
summarizes the travel time estimates. The travel time objects contain,
in total, 97,784,850 origin-destination pairs (observations) from
population to employment locations for all the DAs in the sample and
18,519,897 pairs from population to grocery stores. Considering all
areas combined, the mean travel time to jobs was 78 and 76 to grocery
stores.

\begin{table}
\centering
\caption{\label{tab:unnamed-chunk-1}\label{tab:table_1}Descriptive statistics of the travel time matrices from the {canaccessR} package.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{llrrrrrr}
\toprule
Study Region Name & Destination & Observations & Mean & Sd & P25 & P50 & P75\\
\midrule
\cellcolor{gray!10}{All regions} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{97784850} & \cellcolor{gray!10}{78.06} & \cellcolor{gray!10}{25.18} & \cellcolor{gray!10}{59} & \cellcolor{gray!10}{80} & \cellcolor{gray!10}{99}\\
Toronto & Employment & 41038062 & 81.99 & 24.74 & 64 & 85 & 103\\
\cellcolor{gray!10}{Montréal} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{30091411} & \cellcolor{gray!10}{76.18} & \cellcolor{gray!10}{25.26} & \cellcolor{gray!10}{57} & \cellcolor{gray!10}{77} & \cellcolor{gray!10}{97}\\
Vancouver & Employment & 12254478 & 75.90 & 24.87 & 58 & 77 & 96\\
\cellcolor{gray!10}{Calgary} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{3022244} & \cellcolor{gray!10}{74.57} & \cellcolor{gray!10}{22.60} & \cellcolor{gray!10}{59} & \cellcolor{gray!10}{75} & \cellcolor{gray!10}{91}\\
\addlinespace
Ottawa & Employment & 2936912 & 74.11 & 24.16 & 56 & 74 & 93\\
\cellcolor{gray!10}{Edmonton} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{2057486} & \cellcolor{gray!10}{70.11} & \cellcolor{gray!10}{23.37} & \cellcolor{gray!10}{53} & \cellcolor{gray!10}{70} & \cellcolor{gray!10}{87}\\
Québec City & Employment & 1473702 & 72.46 & 25.37 & 54 & 72 & 93\\
\cellcolor{gray!10}{Winnipeg} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{1574205} & \cellcolor{gray!10}{61.75} & \cellcolor{gray!10}{21.58} & \cellcolor{gray!10}{46} & \cellcolor{gray!10}{61} & \cellcolor{gray!10}{76}\\
Hamilton & Employment & 2089172 & 82.00 & 27.56 & 61 & 87 & 106\\
\addlinespace
\cellcolor{gray!10}{Waterloo} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{552826} & \cellcolor{gray!10}{70.77} & \cellcolor{gray!10}{26.62} & \cellcolor{gray!10}{49} & \cellcolor{gray!10}{69} & \cellcolor{gray!10}{93}\\
London & Employment & 415174 & 60.99 & 20.98 & 46 & 60 & 74\\
\cellcolor{gray!10}{Halifax} & \cellcolor{gray!10}{Employment} & \cellcolor{gray!10}{279178} & \cellcolor{gray!10}{65.96} & \cellcolor{gray!10}{25.45} & \cellcolor{gray!10}{47} & \cellcolor{gray!10}{65} & \cellcolor{gray!10}{84}\\
All regions & Groceries Stores & 18519897 & 75.54 & 25.66 & 56 & 76 & 97\\
\cellcolor{gray!10}{Toronto} & \cellcolor{gray!10}{Groceries Stores} & \cellcolor{gray!10}{8512874} & \cellcolor{gray!10}{79.56} & \cellcolor{gray!10}{25.34} & \cellcolor{gray!10}{61} & \cellcolor{gray!10}{82} & \cellcolor{gray!10}{101}\\
\addlinespace
Montréal & Groceries Stores & 2993965 & 73.14 & 25.68 & 53 & 74 & 94\\
\cellcolor{gray!10}{Vancouver} & \cellcolor{gray!10}{Groceries Stores} & \cellcolor{gray!10}{4540106} & \cellcolor{gray!10}{72.84} & \cellcolor{gray!10}{25.37} & \cellcolor{gray!10}{53} & \cellcolor{gray!10}{74} & \cellcolor{gray!10}{93}\\
Calgary & Groceries Stores & 502757 & 70.03 & 22.25 & 54 & 70 & 86\\
\cellcolor{gray!10}{Ottawa} & \cellcolor{gray!10}{Groceries Stores} & \cellcolor{gray!10}{619791} & \cellcolor{gray!10}{72.19} & \cellcolor{gray!10}{24.41} & \cellcolor{gray!10}{55} & \cellcolor{gray!10}{72} & \cellcolor{gray!10}{91}\\
Edmonton & Groceries Stores & 324030 & 68.07 & 23.41 & 51 & 67 & 84\\
\addlinespace
\cellcolor{gray!10}{Québec City} & \cellcolor{gray!10}{Groceries Stores} & \cellcolor{gray!10}{234600} & \cellcolor{gray!10}{69.66} & \cellcolor{gray!10}{25.79} & \cellcolor{gray!10}{51} & \cellcolor{gray!10}{70} & \cellcolor{gray!10}{89}\\
Winnipeg & Groceries Stores & 362566 & 58.42 & 20.96 & 43 & 57 & 72\\
\cellcolor{gray!10}{Hamilton} & \cellcolor{gray!10}{Groceries Stores} & \cellcolor{gray!10}{281118} & \cellcolor{gray!10}{83.61} & \cellcolor{gray!10}{29.15} & \cellcolor{gray!10}{60} & \cellcolor{gray!10}{91} & \cellcolor{gray!10}{110}\\
Waterloo & Groceries Stores & 44726 & 66.06 & 27.62 & 44 & 62 & 90\\
\cellcolor{gray!10}{London} & \cellcolor{gray!10}{Groceries Stores} & \cellcolor{gray!10}{52617} & \cellcolor{gray!10}{55.48} & \cellcolor{gray!10}{19.26} & \cellcolor{gray!10}{42} & \cellcolor{gray!10}{55} & \cellcolor{gray!10}{67}\\
\addlinespace
Halifax & Groceries Stores & 50747 & 63.16 & 26.39 & 42 & 63 & 83\\
\bottomrule
\end{tabular}}
\end{table}

@@

This section presents a usage example of the package through a visual
representation of its data. In Figure
\ref{fig:fig-travel_time_emp_grc_plot}, we show the spatial distribution
of the travel time matrices for the metropolitan region of Montréal. In
it, we see the median travel time from each DA to employment (left) and
to grocery stores (right). The plot shows that moving away from the city
core increases the necessary travel time by public transit to reach both
employment locations and grocery stores. This result is expected for
employment, given the typical concentration of this type of opportunity
in downtown centers, but it also points out at a distribution of grocery
stores that also requires increased travel for non-central city
residents.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{../figures/patch_tt_emp_grc} \caption{\label{fig:fig-travel_time_emp_grc_plot}Estimated median travel time (minutes) per Dissemination Area to jobs (left) and grocery stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021), and DA and planning boundaries of the Montréal metropolitan area use Statistics Canada data (Statistics Canada, 2016).}\label{fig:unnamed-chunk-2}
\end{figure}

A more thorough example of the package's use can be found in the School
of Cities' recent report on Canada's Urban Infrastructure Deficit. In
its 11th Chapter, the travel time matrices in the package are used to
estimate accessibility metrics to jobs and grocery stores before and
after the pandemic \citep{pargaDemocraticAccessOur2024}. This work also
compares how changes affected groups differently according to their
spatial distribution and income level, thus making explicit the
connection between the package's information and matters of equity in
transportation. The report is freely available for download at the State
of Cities Summit \href{https://stateofcitiessummit.ca/report}{website}.

@@

\bibliographystyle{sageh}
\bibliography{bibfile.bib}
